---
title: " Is there an association between early childhood adversity and higher depression scores in US adults?"
author: "Vincent Yu"
date: "2024-6-18"
output:
  pdf_document:
    latex_engine: xelatex
always_allow_html: true
---

```{r}
#needed questionnaires
#install.packages("haven")
#install.packages("nhanesA")
#install.packages("tableone")
library(nhanesA)
library(haven)
library(tidyverse)
library(tableone)
library(dplyr)
library(tableone)
library(kableExtra)
library(ggplot2)
library(broom)
#depress <- read_xpt("D:/_uoft_1/5硕士1/depress-and-childhood/data/DPQ_J.XPT")
demo <- nhanes("DEMO_J")
depress <- nhanes("DPQ_J")
#child <- read_xpt("D:/_uoft_1/5硕士1/depress-and-childhood/data/ECQ_J.XPT")
child <- nhanes("ECQ_J")
BMI <- nhanes("BMX_J")
```
The relationship between early childhood adversity and higher depression scores is complex and multifaceted. 
Women with a birth weight of less than 3 kg had an increased risk of depression at age 26 compared to those with a birth weight of more than 3.5 kg.
Older maternal age is a predictor of adverse mental health symptoms, particularly in female offspring, suggesting a gender-specific effect.

Early exposure to significant stressors can lead to long-term changes in brain function and increase the risk of developing depressive disorders. Women appear to be more affected by these adversities, reporting higher rates of depression than men. Specific stressors such as parental divorce, frightening events, and physical abuse significantly increase the likelihood of depression. Parental influence, both genetic and environmental, plays a crucial role in shaping children's health outcomes, further emphasizing the need for supportive and stable family environments.

Based on these findings and considering the limited data available, I have formulated the following research questions for further investigation:
How the participants was affected
H1: Lower birth weight is associated with higher depression scores. ECD070A
H2: Older maternal age at birth is associated with higher depression scores, particularly in female participants. ECD010 
H3: Maternal smoking during pregnancy is associated with higher depression scores. ECQ020

And how these will affect the next generation
H4: Higher BMI in participants is associated with higher depression scores. 
H5: Higher depression scores in participants are associated with adverse health outcomes in the next generation. MCQ080E
I am not sure if this is the right direction and would greatly appreciate any guidance or suggestions you could provide.
```{r}
# merge 2 datasets
depress <- nhanes("DPQ_J")%>% filter_all(all_vars(!is.na(.)))
# Replace value descriptions with code values
depress_replaced <- depress %>%
  mutate(across(starts_with("DPQ"), ~ recode(.,
                                             "Not at all" = 0,
                                             "Several days" = 1,
                                             "More than half the days" = 2,
                                             "Nearly every day" = 3)))
# Replace DPQ100 with the sum of DPQ010 to DPQ090
depress_replaced <- depress_replaced %>%
  mutate(DPQ100 = rowSums(select(., starts_with("DPQ0"))))

clean_data <- merge(demo, depress_replaced,by = "SEQN", all.x =T)
clean_data_T <- merge(clean_data, child, by = "SEQN", all.x =T)
clean_data_T <- merge(clean_data_T, BMI, by = "SEQN", all.x =T)
# filter missing and refused

```

#filter missing value still error: ModuleStopIfNoVarsLeft(vars): No valid variables. e.x how to handle missing value. none of the "SEQN" is matched 
```{r}
clean_data_F <- merge(depress, BMI, by = "SEQN", all.x =F, all.y =F)
clean_data_FF <- merge(depress, child, by = "SEQN", all.x =F, all.y =F)
```

```{r, warning=FALSE}
demo_clean <- clean_data_T %>% select("SEQN","DPQ100","RIDAGEYR","DMDHRGND","RIDRETH1","ECD070A","ECD010","ECQ020", "DMDMARTL","BMXBMI","MCQ080E") 
colnames(demo_clean) <- c("SEQN","Depress","Age", "Gender", "Race","birth_weight","maternal_age","Maternal_smoking", "Marital_status", "BMI", "adverse health outcomes in the next generation")
# Define variables, factorVars for categorical variable
vars <- c("Age", "Gender", "Race","birth_weight","maternal_age","Maternal_smoking", "Marital_status", "BMI", "adverse health outcomes in the next generation")
factorVars <- c("Gender", "Race", "Maternal_Smoking", "Marital_Status")
table1 <- CreateTableOne(vars = vars, data = demo_clean, factorVars = factorVars)


# Convert to data frame for stargazer
table1_df <- as.data.frame(print(table1, printToggle = FALSE))
# Display the table in R
kable(table1_df, caption = "table.1 Demographics")
```
# what analysis model should I use


```{r}
#model1 <- glm(DPQ100 ~ Maternal_smoking + Age + Gender + Race + birth_weight, data = demo_clean, family = binomial)
#model2 <- glm(DPQ100 ~ Maternal_smoking + Age + Gender + Race + maternal_age, data = demo_clean, family = binomial)
#model3 <- glm(DPQ100 ~ Maternal_smoking + Age + Gender + Race + Marital_status , data = demo_clean, family = binomial)
#model4 <- glm(DPQ100 ~ Maternal_smoking + Age + Gender + Race + BMI + adverse_health_outcomes_next_generation, data = demo_clean, family = binomial)
# Summary of the model
#summary(model1)
```
```{r}
# Create quantile groups for BMI
demo_clean_BMI <- demo_clean %>%
  mutate(BMI_quantiles = ntile(BMI, 4)) %>%
  mutate(BMI_quantiles = factor(BMI_quantiles, labels = c("Q1", "Q2", "Q3", "Q4")))
# Create Table 1 stratified by BMI quantiles
table2 <- CreateTableOne(vars = vars, strata = "BMI_quantiles", data = demo_clean_BMI, factorVars = factorVars)

# Convert to data frame for kableExtra
table2_df <- as.data.frame(print(table2, printToggle = FALSE))

# Use kableExtra for better formatting
kable(table2_df, caption = "Demographics") 

```
```{r}
model1 <- glm(Depress ~ BMI, data = demo_clean)
summary(model1)
```
```{r}
data <- demo_clean %>% filter(!is.na(Depress))%>%
  mutate(Depress = rowSums(select(., starts_with("Depress"))))

# Calculate percentage of depressive symptoms for each BMI integer category
percentage_data <- data %>%
  group_by(BMI) %>%
  summarise(depress_count = sum(Depress),
            total_count = n(),
            percentage_depress = (depress_count / total_count)) 

# Create the plot
ggplot(percentage_data, aes(x = BMI, y = percentage_depress)) +
  geom_smooth(color = "blue", method = "loess", span = 0.3) +
  labs(title = "Percentage of Depressive Symptoms by BMI",
       x = "BMI",
       y = "Percentage of Depressive Symptoms") +
  theme_minimal() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
```
```{r}
# Create the density plot for BMI
ggplot(demo_clean, aes(x = BMI)) +
  geom_density(fill = "blue", alpha = 0.3) +
  labs(title = "Density Plot of BMI", x = "BMI", y = "Density") +
  theme_minimal() +
  theme(axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5))

```









